---
title: 'PSI CA PART I '
R Packages: dplyr, pastecs, ggplot2, pysch, semTools, car, pander
R Version: 3.5.1
output: 
  pdf_document: default
  html_document: default
Student Number: D17129566
Student Name: PRIYAM BANSAL
Programme Code: DT228A
---

#Introduction

#1.1 Research Question

A person's overall health can be impacted by a range of social, pyschological and physical factors. This analysis is
concerned with a person's smoking habbits, sleep impact on energy level, sleep impact on mood , restless sleepers,
anxiety, sleep impact on overall well-being, caffeine consumption each day, general health rate, depression, and the
satisfaction in the amount of sleep.
when overall health is considered then evergy level and sleep factors are also a part in explaining that as in this
cigarettes smoked per day can determine the energy level thatsleep has an impact on , another factor that sllep can
impact on is mood which can have some relatin with restless sleep which will directly effect on the health next
considering anxiety issues when a person will be having anxiety and its level is higher it it effect its sleep imact
on overall well-being will will consicuently effect on overall health. there can be a relation or a difference
between caffeine cosumption each day and its effect on general health. Depression leads to a huge part in the health
a person is having which will effect the satisfaction in the amount of sleep which will lead to inspect the overall
health.The hypotheses made below are looked for as there is actually relationship or difference between the variales?


##1.2 Hypotheses

The particular hypotheses to be investigated are:

####Hypothesis 1:

H0:  There is no correlation between cigarettes smoked per day and Rate of impact of sleep on energy level

HA:  There is a correlation between cigarettes smoked per day and Rate of impact of sleep  on energy level

####Hypothesis 2:

H0:There is no difference in sleep impact on mood for the restless sleeper or not a restless sleeper

HA:There is a difference in sleep impact on mood for the restless sleeper or not a restless sleeper

####Hypothesis 3:

H0:There is no correlation between level of anxiety and sleep impact on overall well-being

HA:There is a correlation between level of anxiety and sleep impact on overall well-being

####Hypothesis 4:

H0:There is no difference in drinks containing caffeine per day or general health rate

HA:There is a difference in drinks containing caffeine per day or general health rate

####Hypothesis 5:

H0:There is no difference in level of depression or satisfaction with the amount of sleep

HA:There is a difference in level of depression or satisfaction with the amount of sleep

```{r set_up}
library(rmarkdown)
library(effsize)
library(lsr)
library(ggplot2) #For creating histograms with more detail than plot
library(psych) # Some useful descriptive functions
library(pastecs)#Some useful descriptive functions
library(lavaan)
library(semTools)#For skewness and kurtosis
library(carData)
library(car) # For Levene's test for homogeneity of variance 
library(pander) # For creating a table
library(dplyr)
library(FSA)
library(dunn.test)


#Read in the dataset
sleep <- read.csv('sleep5ED.csv', header = TRUE, sep = ",")
View(sleep)
```
#2. METHODOLOGY

##2.1 Variables of interest
The variables of interest used in this research are shown below:
  
```{r var_interest}
my.data <- " 
  Concept                                      | Variable Name    | Statistical Type      | Possible Values
  How many cigarettes per day                    | somkenum        | scale                  | Cigs per day 
  Rate impact of sleep problem on energy level   | impact2         | Scale                  | 1=not at all 10=to a great extent 
  
  Restless sleeper                               | restlss         | Nominal                | 1=yes 2=no
  Rate impact of sleep problem on mood           | impact1         | Scale                  | 1=not at all 10=to a great extent
   HADS Anxiety                                  | anxiety         | Scale                  | range from 0=no anxiety to 21=severe anxiety 
 Rate impact of sleep problem on overall         | impact6         | Scale                  | 1=not at all 10=to a great extent 
  well-being 
  How many caffeine drinks per day               | caffeine        |Scale                   |drink per day
  Rate general health                            | healthrate      | scale                  | 1=very poor 10=very good 
   HADS Depression                               |depress          | scale                  |range from 0=no depression to 21=severe depression 
  Satisfaction with amount of sleep              | satsleep        | Scale                  | 1=very dissatisfied 10=very satisfied "

df <- read.delim(textConnection(my.data),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,])) # headers
df <- df[-1,] # remove first row
row.names(df)<-NULL
pander(df, style = 'rmarkdown')
View(df)
```

##2.2 cigarettes smoked per day 

###Code for inspecting the variable

```{r inspect_var1}
pastecs::stat.desc(sleep$smokenum, basic=F)
#skewness and kurtosis from semTools with standard error 
cgskew<-semTools::skew(sleep$smokenum)
cgkurt<-semTools::kurtosis(sleep$smokenum)
cgskew
cgkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew1<-cgskew[1]/cgskew[2]
stdkurt1<-cgkurt[1]/cgkurt[2]
stdkurt1
stdskew1

outliers <- sleep %>% 
  filter(scale(smokenum) >3.29 | scale(smokenum) < -3.29)
outliers

#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis

Cigarettes smoked per day is represented by a scale variable in the dataset which was calculated from the Overall
Health  questionnaire. Inspection of the histogram and normality plot (see figure 1 and figure 2) shows that the
distribution does not conform exactly to a normal distribution.
Inspection of standardized normal scores for skewness and kurtosis indicated that while both do not fall with an
acceptable range of +/-2, (skewness of `r cgskew[1]` SE=`r cgskew[2]`, kurtosis of `r cgkurt[1]` SE=`r cgkurt[2]`.
Further inspection of the variable using standardized scores showed that `r round((numoutliers/fullcount)*100,
digits=2)`% standardised scores were inside  the acceptable range of +/3.29. Total self esteem will therefore be
treated as a not normal within this analysis (stat.desc(sleep$smokenum, basic=F))

```{r plots_var1}
ggplot(sleep, aes(x = smokenum), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow') + 
  geom_density(colour = 'red') + xlab(expression(bold('cigarette smoked per day'))) + 
  ylab(expression(bold('Density')))+ggtitle('Figure 1: Number of cigarettes smoked per day')

#Create a qqplot
qqnorm(sleep$smokenum,  main="Figure 2: QQPlot or cigarettes smoked per day")
qqline(sleep$smokenum, col=2)
```
## 2.3 RATE OF IMPACT OF SLEEP ON ENERGY LEVEL

###Code for inspecting the variable

```{r inspect_var2}
pastecs::stat.desc(sleep$impact2, basic=F)
#skewness and kurtosis from semTools with standard error 
elskew<-semTools::skew(sleep$impact2)
elkurt<-semTools::kurtosis(sleep$impact2)
elskew
elkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew2<-elskew[1]/elskew[2]
stdkurt2<-elkurt[1]/elkurt[2]
stdkurt2
stdskew2

outliers <- sleep %>% 
  filter(scale(impact2) >3.29 | scale(impact2) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis

sleep impact on energy level is represented by a scale variable in the dataset which was calculated from the Overall
Health  questionnaire. Inspection of the histogram and normality plot (see figure 3 and figure 4) shows that the
distribution does not conform exactly to a normal distribution.
Inspection of standardized normal scores for skewness and kurtosis indicated that while both do not fall with an
acceptable range of +/-2, (skewness of `r elskew[1]` SE=`r elskew[2]`, kurtosis of `r elkurt[1]` SE=`r elkurt[2]`.
Further inspection of the variable using standardized scores showed that `r round((numoutliers/fullcount)*100,
digits=2)`% standardised scores were inside  the acceptable range of +/3.29. Total self esteem will therefore be
treated as a not normal within this analysis (stat.desc(sleep$impact2, basic=F))

```{r plot_var2 }
ggplot(sleep, aes(x = impact2), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('sleep impact on energy level'))) + 
  ylab(expression(bold('Density')))+ ggtitle("Figure3:Rate of impact of sleep on Energy Level ")

qqnorm(sleep$impact2, main = "Figure3: QQPlot or sleep impact on energy level")
qqline(sleep$impact2, col=2)
```
##2.4 RESTLESS SLLEPER

restless sleeper is represented by a categorical variable in the dataset. The dataset contains data from 88 yes
respondents and 176 no respondents.
```{r somdesc_restlss}
table(sleep$restlss)

#Get descriptive stastitics by group
psych::describeBy(sleep$impact1, sleep$restlss)
```

```{r bar_restlss}
bar <- ggplot(sleep, aes(restlss, impact1))
bar<- bar + stat_summary(fun.y = mean, geom = "bar", fill = "black", colour = "white", na.rm = FALSE) +
  stat_summary(fun.data = mean_cl_normal, geom = "pointrange", na.rm = FALSE) + labs(x = "restless sleep", y = "
 impact of sleep problem on mood", title="Figure 5:Rate of impact of sleep problem on mood by restless sleep") 

bar
```

## 2.5 RATE OF IMPACT OF SLEEP ON MOOD

###Code for inspecting the variable

```{r inspect_var4}
pastecs::stat.desc(sleep$impact1, basic=F)
#skewness and kurtosis from semTools with standard error 
mdskew<-semTools::skew(sleep$impact1)
mdkurt<-semTools::kurtosis(sleep$impact1)
mdskew
mdkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew4<-mdskew[1]/mdskew[2]
stdkurt4<-mdkurt[1]/mdkurt[2]
stdkurt4
stdskew4

outliers <- sleep %>% 
  filter(scale(impact1) >3.29 | scale(impact1) < -3.29)

#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```
###Report of normality analysis
Rate of impact of sleep on mood is represented by a scale variable in the dataset which was calculated from the Total
Perceived Control questionnaire. Inspection of the histogram and normality plot (see figure 5 and figure 6) shows
that the distribution does not conform exactly to a normal distribution. Inspection of standardized normal scores for
skewness and kurtosis indicated that while skewness was not an issue  falling with an acceptable range of +/-2,
(skewness of `r mdskew[1]` SE=`r mdskew[2]`), kurtosis was outside acceptable bounds, (kurtosis of `r mdkurt[1]`
SE=`r mdkurt[2]` . Further inspection of the variable using standardized scores showed that only `r numoutliers` or
`r round((numoutliers/fullcount)*100, digits=2)`% standardised scores were outside the acceptable range of +/3.29.
Total perceived control of internal states will therefore be treated as a normal within this analysis
pastecs::stat.desc(sleep$impact1, basic=F)

```{r polt_var4}
ggplot(sleep, aes(x = impact1), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('sleep impact on mood'))) + 
  ylab(expression(bold('Density'))) + ggtitle("Figure5: Rate of impact of sleep on mood")
qqnorm(sleep$impact1, main = "Figure6 :QQ plot or sleep impact on mood")
qqline(sleep$impact1, col=2)
```

## 2.6 LEVEL OF ANXIETY

###Code for inspecting the variable
```{r inspect_var5}
pastecs::stat.desc(sleep$anxiety, basic=F)
#skewness and kurtosis from semTools with standard error 
anskew<-semTools::skew(sleep$anxiety)
ankurt<-semTools::kurtosis(sleep$anxiety)
ankurt
anskew

#We divide the skew statistic by the standard error to get the standardised score
stdskew5<-anskew[1]/anskew[2]
stdkurt5<-ankurt[1]/ankurt[2]
stdskew5
stdkurt5

outliers <- sleep %>% 
  filter(scale(anxiety) >3.29 | scale(anxiety) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis
Level of anxiety is represented by a scale variable in the dataset which was calculated from the Total Perceived
Control questionnaire. Inspection of the histogram and normality plot (see figure 7 and figure 8) shows that the
distribution does not conform exactly to a normal distribution. Inspection of standardized normal scores for skewness
and kurtosis indicated that while skewness was an issue  falling with an acceptable range of +/-2, (skewness of `r
anskew[1]` SE=`r anskew[2]`), kurtosis was not outside acceptable bounds, (kurtosis of `r ankurt[1]` SE=`r ankurt[2]`
. Further inspection of the variable using standardized scores showed that only `r numoutliers` or `r
round((numoutliers/fullcount)*100, digits=2)`% standardised scores were not outside the acceptable range of +/3.29.
Total perceived control of internal states will therefore be treated as not normal within this analysis
pastecs::stat.desc(sleep$anxiety, basic=F)

```{r plot_var5}
ggplot(sleep, aes(x = anxiety), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('anxiety'))) + 
  ylab(expression(bold('Density'))) + ggtitle("Figure7: Level of Anxiety ")
qqnorm(sleep$anxiety, main = "Figure8: QQPlot or level of anxiety")
qqline(sleep$anxiety, col=2)
```

## 2.7 SLEEP IMPACT ON OVERALL WELL-BEING

###Code for inspecting the variable
```{r inspect_var6}
pastecs::stat.desc(sleep$impact6, basic=F)
#skewness and kurtosis from semTools with standard error 
obskew<-semTools::skew(sleep$impact6)
obkurt<-semTools::kurtosis(sleep$impact6)
obskew
obkurt
#We divide the skew statistic by the standard error to get the standardised score
stdskew6<-obskew[1]/obskew[2]
stdkurt6<-obkurt[1]/obkurt[2]
stdkurt6
stdskew6

outliers <- sleep %>% 
  filter(scale(impact6) >3.29 | scale(impact6) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis
Rate of impact of sleep on overall well-being is represented by a scale variable in the dataset which was calculated
from the Total Perceived Control questionnaire. Inspection of the histogram and normality plot (see figure 9 and
figure10) shows that the distribution does not conform exactly to a normal distribution. Inspection of standardized
normal scores for skewness and kurtosis indicated that while skewness was not an issue  falling with an acceptable
range of +/-2, (skewness of `r obskew[1]` SE=`r obskew[2]`), kurtosis was not outside acceptable bounds, (kurtosis of
`r obkurt[1]` SE=`r obkurt[2]` . Further inspection of the variable using standardized scores showed that only `r
numoutliers` or `r round((numoutliers/fullcount)*100, digits=2)`% standardised scores were not outside the acceptable
range of +/3.29. Total perceived control of internal states will therefore be treated as a normal within this
analysis pastecs::stat.desc(sleep$impac6, basic=F)
```{r plot_var6}
ggplot(sleep, aes(x = impact6), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('sleep impact on overall wellbeing'))) + 
  ylab(expression(bold('Density'))) + ggtitle("Figure9: Rate of impact of sleep on overall well_being")
qqnorm(sleep$impact6, main = "Figure 10:QQPlot or sleep impact on overall well-being")
qqline(sleep$impact6, col=2)
```

## 2.8 DRINKS CONTAINING CAFFEINE PER DAY

###Code for inspecting the variable
```{r inspect_var7}
pastecs::stat.desc(sleep$caffeine, basic=F)
#skewness and kurtosis from semTools with standard error 
cfskew<-semTools::skew(sleep$caffeine)
cfkurt<-semTools::kurtosis(sleep$caffeine)
cfskew
cfkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew7<-cfskew[1]/cfskew[2]
stdkurt7<-cfkurt[1]/cfkurt[2]
stdkurt7
stdskew7

outliers <- sleep %>% 
  filter(scale(caffeine) >3.29 | scale(caffeine) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```
###Report of normality analysis
Drinks containing caffeine per day is represented by a scale variable in the dataset which was calculated from the
Total Perceived Control questionnaire. Inspection of the histogram and normality plot (see figure 11 and figure 12)
shows that the distribution does not conform exactly to a normal distribution. Inspection of standardized normal
scores for skewness and kurtosis indicated that while skewness was an issue  falling with an acceptable range of
+/-2, (skewness of `r cfskew[1]` SE=`r cfskew[2]`), kurtosis was not outside acceptable bounds, (kurtosis of `r
cfkurt[1]` SE=`r cfkurt[2]` . Further inspection of the variable using standardized scores showed that only `r
numoutliers` or `r round((numoutliers/fullcount)*100, digits=2)`% standardised scores were not outside the acceptable
range of +/3.29. Total perceived control of internal states will therefore be treated as not normal within this
analysis pastecs::stat.desc(sleep$caffeine, basic=F)

```{r plot_var7}
ggplot(sleep, aes(x = caffeine), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('caffeine per day'))) + 
  ylab(expression(bold('Density'))) + ggtitle("Figure11: drinks containing caffeine per day")
qqnorm(sleep$caffeine, main="Figure12: QQPlot or drinks containing caffeine per day")
qqline(sleep$caffeine, col=2)
```

## 2.9 GENERAL HEALTH RATE

###Code for inspecting the variable
```{r inspect_var8}
pastecs::stat.desc(sleep$healthrate, basic=F)
#skewness and kurtosis from semTools with standard error 
hrskew<-semTools::skew(sleep$healthrate)
hrkurt<-semTools::kurtosis(sleep$healthrate)
hrskew
hrkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew8<-hrskew[1]/hrskew[2]
stdkurt8<-hrkurt[1]/hrkurt[2]
stdkurt8
stdskew8

outliers <- sleep %>% 
  filter(scale(healthrate) >3.29 | scale(healthrate) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis
General health rate is represented by a scale variable in the dataset which was calculated from the Total Perceived
Control questionnaire. Inspection of the histogram and normality plot (see figure 13 and figure 14) shows that the
distribution does not conform exactly to a normal distribution. Inspection of standardized normal scores for skewness
and kurtosis indicated that while skewness was an issue  falling with an acceptable range of +/-2, (skewness of `r
hrskew[1]` SE=`r hrskew[2]`), kurtosis was not outside acceptable bounds, (kurtosis of `r hrkurt[1]` SE=`r hrkurt[2]`
. Further inspection of the variable using standardized scores showed that only `r numoutliers` or `r
round((numoutliers/fullcount)*100, digits=2)`% standardised scores were not outside the acceptable range of +/3.29.
Total perceived control of internal states will therefore be treated as not normal within this analysis
pastecs::stat.desc(sleep$healthrate, basic=F)

```{r plot_var8}
ggplot(sleep, aes(x = healthrate), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('general health'))) + 
  ylab(expression(bold('Density')))+ggtitle("Figure13 : Rate of General Health")
qqnorm(sleep$healthrate, main="Figure13 : QQPlot or Rate of General Health")
qqline(sleep$healthrate, col=2)
```

## 2.10 LEVEL OF DEPRESSION

###Code for inspecting the variable
```{r inspect_var9}
pastecs::stat.desc(sleep$depress, basic=F)
#skewness and kurtosis from semTools with standard error 
dpskew<-semTools::skew(sleep$depress)
dpkurt<-semTools::kurtosis(sleep$depress)
dpskew
dpkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew9<-dpskew[1]/dpskew[2]
stdkurt9<-dpkurt[1]/dpkurt[2]
stdkurt9
stdskew9

outliers <- sleep %>% 
  filter(scale(depress) >3.29 | scale(depress) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis
Level of depression is represented by a scale variable in the dataset which was calculated from the Total Perceived
Control questionnaire. Inspection of the histogram and normality plot (see figure 15 and figure 16) shows that the
distribution does not conform exactly to a normal distribution. Inspection of standardized normal scores for skewness
and kurtosis indicated that while skewness was an issue  falling with an acceptable range of +/-2, (skewness of `r
dpskew[1]` SE=`r dpskew[2]`), kurtosis was not outside acceptable bounds, (kurtosis of `r dpkurt[1]` SE=`r dpkurt[2]`
. Further inspection of the variable using standardized scores showed that only `r numoutliers` or `r
round((numoutliers/fullcount)*100, digits=2)`% standardised scores were not outside the acceptable range of +/3.29.
Total perceived control of internal states will therefore be treated as not normal within this analysis
pastecs::stat.desc(sleep$depress, basic=F)

```{r plot_var9}
ggplot(sleep, aes(x = depress), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('depression'))) + 
  ylab(expression(bold('Density')))+ ggtitle("Figure15 :Level of Depression")
qqnorm(sleep$depress, main="Figure15 :QQPlot or  Level of Depression")
qqline(sleep$depress, col=2)
```

## 2.11 SATISFATION WITH THE AMOUNT OF SLEEP

###Code for inspecting the variable
```{r inspect_var10}
pastecs::stat.desc(sleep$satsleep, basic=F)
#skewness and kurtosis from semTools with standard error 
stskew<-semTools::skew(sleep$satsleep)
stkurt<-semTools::kurtosis(sleep$satsleep)
stskew
stkurt

#We divide the skew statistic by the standard error to get the standardised score
stdskew10<-stskew[1]/stskew[2]
stdkurt10<-stkurt[1]/stkurt[2]
stdkurt10
stdskew10

outliers <- sleep %>% 
  filter(scale(satsleep) >3.29 | scale(satsleep) < -3.29)
outliers
#count them using dplyr summarize
numoutliers<-outliers %>%
  summarize(count=n())
fullcount<-sleep %>%
  summarize(count=n())
numoutliers
fullcount
```

###Report of normality analysis
Satisfaction with the amount of sleep is represented by a scale variable in the dataset which was calculated from the
Total Perceived Control questionnaire. Inspection of the histogram and normality plot (see figure 17 and figure18)
shows that the distribution does not conform exactly to a normal distribution. Inspection of standardized normal
scores for skewness and kurtosis indicated that while skewness was not an issue  falling with an acceptable range of
+/-2, (skewness of `r stskew[1]` SE=`r stskew[2]`), kurtosis was outside acceptable bounds, (kurtosis of `r
stkurt[1]` SE=`r stkurt[2]` . Further inspection of the variable using standardized scores showed that only `r
numoutliers` or `r round((numoutliers/fullcount)*100, digits=2)`% standardised scores were outside the acceptable
range of +/3.29. Total perceived control of internal states will therefore be treated as a normal within this
analysis pastecs::stat.desc(sleep$impac6, basic=F)

```{r plot_var10}
ggplot(sleep, aes(x = satsleep), binwidth = 2) + 
  geom_histogram(aes(y = ..density..), fill = 'yellow', alpha = 0.5) + 
  geom_density(colour = 'red') + xlab(expression(bold('satisfaction with the amount of sleep'))) + 
  ylab(expression(bold('Density')))+ ggtitle("Figure 17:Satisfaction with the amount of sleep")
qqnorm(sleep$satsleep, main="Figure 18: QQPlot or Satisfaction with the amount of sleep")
qqline(sleep$satsleep, col=2)
```
##1.1 Hypotheses

The particular hypotheses to be investigated are:

####Hypothesis 1:

H0:  There is no correlation between cigarettes smoked per day and Rate of impact of sleep on energy level

HA:  There is a correlation between cigarettes smoked per day and Rate of impact of sleep  on energy level



```{r correlation_cigarettes to energy level}
cor.test(sleep$smokenum, sleep$impact2, method = "spearman")
#spearman correlation
cor.test(sleep$smokenum, sleep$impact2, method = "kendall")
#kendall correlation to calculate accute p-value
scatter<-ggplot(sleep, aes(sleep$smokenum, sleep$impact2))
scatter + geom_point()+ geom_smooth(method = 'lm', colour='Red', se=F)+ labs(x="cigarettes smoked per day", y="rate of impact of sleep on energy level ")
```

The relationship between cigarettes smoked per day and rate impact of sleep on energy level was investigated using a
Spearman correlation because ther was no normality so used a non-parametric Test . A small/weak and negative
correlation was found (rho =-0.1442013 S = 1108.7, p-value = 0.6442).There is evidence to support rejecting the null
hypothesis that there is relationship between cigarettes smoked per day and rate impact of sleep on energy level 


####Hypothesis 2:

H0:There is no difference in sleep impact on mood for the restless sleeper or not a restless sleeper

HA:There is a difference in sleep impact on mood for the restless sleeper or not a restless sleeper

#### Independent t-test
```{r difference restless sleep}

```
A Levene's test for homogeneity of variance was conducted and indicated equality of variance for sleep impact on mood
for restless sleeper(group1) or not restless sleeper(group2) respondents (F=7.468, Pr=0.007). An independent-samples
t-test was conducted to compare sleep impact on mood for group1 and group2 respondents. No significant difference in
the scores for  was found (M=6.07, SD= 2.19 for group respondents, M= 5.4, SD= 2.64 for group 2 respondents), (t=
1.5077, p = 0.1343). The cohen.d test result (2.877(large)) While the result is statistical significance which
indicates there is evidence to accept the null hypothesis that there is no difference in self esteem scores for
respondents of different groups,  the very large effect size is a not a  concern .


####Hypothesis 3:

H0:There is no correlation between level of anxiety and sleep impact on overall well-being

HA:There is a correlation between level of anxiety and sleep impact on overall well-being

```{r correlation_amxiety and well-being}
cor.test(sleep$anxiety, sleep$impact6, method = "spearman")
#spearman correlation
cor.test(sleep$smokenum, sleep$impact6, method = "kendall")
#kendall correlation to calculate accute p-value
scatter<-ggplot(sleep, aes(sleep$anxiety, sleep$impact6))
scatter + geom_point()+ geom_smooth(method = 'lm', colour='Red', se=F)+ labs(x="anxiety", y="sleep impact on overall well-being ")
```

The relationship between level of anxiety and sleep impact on overall well-being was investigated using a Spearman
correlation because ther was no normality so used a non-parametric Test . A medium/moderate and positive correlation
was found (  rho =0.3482 S = 192420, p-value = 0.3687).  There is evidence to support rejecting the null hypothesis
that there is relationship between  

####Hypothesis 4:

H0:There is no difference in drinks containing caffeine per day or general health rate

HA:There is a difference in drinks containing caffeine per day or general health rate

####kruskal-wallis
```{r differnce_on healthrate}
describeBy(sleep$caffeine, sleep$healthrate)

kt<- kruskal.test(caffeine ~ healthrate, data=sleep)
tmp<- dunnTest(x=sleep$caffeine, g=factor(sleep$healthrate), method="bonferroni") 
kt<- as.numeric(kt)
print(tmp, dunn.test.results = TRUE)
n<- sum(table(sleep$caffeine, sleep$healthrate))
n<-as.numeric(n)
#calculated epsilon square forkruskal-wallis
epsilonSq<- kt*(n+1)/(n^2-1) 
epsilonSq
```
A kruskal-wallis test for homogeneity of variance was conducted and indicated equality of variance for drinks
containing caffeine per day for general health rate respondents (chi-square=12.395, df=7 and p-value=0.09) col mean
-row mean(0.320-1.00).there is epsilon squared epsilonSq<- kt*(n+1)/(n^2-1) value for the effect i.e(0.0482) While
the result is statistical significance which indicates there is evidence to reject the null hypothesis that there is
no difference  and p <= alpha.

####Hypothesis 5:

H0:There is no difference in level of depression or satisfaction with the amount of sleep

HA:There is a difference in level of depression or satisfaction with the amount of sleep

####kruskal-wallis
```{r difference}
describeBy(sleep$depress, sleep$satsleep)

kw<- kruskal.test(depress ~ satsleep, data=sleep)
tmp<- dunnTest(x=sleep$depress, g=factor(sleep$satsleep), method="bonferroni") 
kw<- as.numeric(kw)
print(tmp, dunn.test.results = TRUE)
n<- sum(table(sleep$depress, sleep$satsleep))
n<-as.numeric(n)
#calculated epsilon square forkruskal-wallis
epsilonSq<- kw *(n+1)/(n^2-1) 
epsilonSq
``` 
A kruskal-wallis test for homogeneity of variance was conducted and indicated equality of variance for level of
depression and satisfaction in the amount of sleep respondents (chi-square=16.44, df=9 and p-value=0.06) col mean
-row mean(0.201-1.00).there is epsilon squared epsilonSq<- kt*(n+1)/(n^2-1) value for the effect i.e(0.6204) While
the result is statistical significance which indicates there is evidence to reject the null hypothesis that there is
no difference and p <= alpha.

#4.0 DISCUSSION

while comparing cigarettes smoked per day and the sleep impact on energy level there is a weak and a negative
correlation between the two which shows that there is a relation between the two of them but it is weak or say not
that approppraite but then also we can say there is a relation between the two
when considering restless sleepers it has two groups people who arerestless sleepers and those who aren't so when
taking a difference of restless sleepers with the sleep impact on mood we can get a difference and a large difference
due to which the result can be taken out in the normal way . The t-test value says that there is a large value for
the effect
while comparing level of anxiety in people and the sleep impact on overall well-being so while seeing the correlation
between the two we see that they are not normal so we did spearman test on them which gave us a moderate and a
positive relationship between the two there are found many people who have anxiety but there sleepis not that much
effected and visa versa also .so the value we obtained rho is also states that while comparing both the data the
result obtained is moderate
Another hypothesis having the variables people drinking caffeine per day and impact on general health while these data are not categorical and are parametric also we did a kruskal-wallis test for finding the difference , as we see that there is epsilon square done to find the effect which occurs 0.482 . rejecting the null hypothses is the p<= alpha so as here p is 0.09 and alpha is 0.05.
Another hypothesis having the variables level of depression and satisfaction in the amount of sleep while these data
are not categorical and are parametric also we did a kruskal-wallis test for finding the difference , as we see that
there is epsilon square done to find the effect which occurs 0.620 . rejecting the null hypothses is the p<= alpha so
as here p is 0.06 and alpha is 0.05.

#5.0 REFERENCE

Field, A., Miles, J., & Field, Z. (2012). Discovering statistics using R. Sage publications.
George, D. (2011).

